---
title: "Introduction to betticurves packeges"
author: "Rafael Gonzalez Gouveia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introducing betticurves package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

The betticurves packege implements betti curves from topological data analysis. It computes the curve based on a homology diagram. It accepts diagrams from TDA packege and TDAstat packege.

# What is the objective behind the package?

My intention is to refactor all the code I used during my master thesis at CIMAT (Guanajuato, Mexico) and at the same time learn how to put it in a R package and upload it in GitHub. In my thesis I analyse how betti curve can be implemented in a data science framework. My work can be seen on [This link](https://gonzalezgouveia.github.io/documents/TesisMaestriaRafael.pdf). Unfortunatelly by the time this documentation is been written (jan 2019) my work is only available in spanish. I am working on a English draft. You can also contact me via Git Hub if you are more interested on my work.

# How to install the package from github?

By now the package is not available in CRAN. So in order to use it in your R session you can install first the `devtools` packages (documentation [here](https://www.rstudio.com/products/rpackages/devtools/)) and then use the following code:
```{r, eval=FALSE}
# install package from GitHub
devtools::install_github("gonzalezgouveia/betticurves", force=TRUE, build_opts = c("--no-resave-data", "--no-manual"))
```
The github repo link [here](https://github.com/gonzalezgouveia/betticurves/). This version may be unstable and it is under development.

# How to use betticurves?

By now, the packege only have two functions for computations and one for visualization:

* The function `compute_betti_number` is for Betti number of a persistence diagram for a desired dimension and radius.
* The function `compute_betti_curve` Compute Betti curves of a persistence diagram until a desired dimension, a maximum radius, and output vector lenght. So betti_number is been used inside betti_curve function. 
* The `compute_betti_curve` will return a tibble which is ready to be plotted using the `plot_betti_curve` function. 

## Example workflow

Let's take a look at an example of the basic workflow. I will asume you have already downloaded the package from GitHub. This example uses the function `calculate_homology` from the TDAstats package (documentation [here](https://github.com/rrrlw/TDAstats)). You can install this package directly from CRAN.

So first we load the packages in the R session
```{r}
library("betticurves")
library("TDAstats")
```

In this example we will the data from a two dimentional uniform point process. So it will have 100 uniformly distributed random points for the first coordinate and the same for the second coordinate. 

```{r}
set.seed(12321)
data <- cbind(runif(100), runif(100))
plot(data, main='Data for the example')
```

Next, we calculate the homology diagram. Here I decided to use the `TDAstat` package because it is faster due to an implementation of Ripser algorithm, but it works fine with diagrams from `TDA` packages(`ripsDiag` or `alphaComplexDiag`), just take a look at the commented line to know how to extract the matrix from the 'diagram' object of TDA.

```{r}
diag <- TDAstats::calculate_homology(data)            # for TDAstat package
# diag <- TDA::alphaComplexDiag(data)$diagram    # for TDA package
betti_curve <- betticurves::compute_betti_curves(diag)
print(summary(betti_curve))
```

The betti_curve tibble has three variables: 

* `dim`: the dimensions of the homology. Here only 0 and 1.
* `radius`: the radius of the complex filtration. Remember points are generated in the interval [0,1].
* `value`: the betti number for a radius. We have circa 100 points.

We can see the resulting curves with the function `plot_betti_curve`

```{r}
plot_betti_curve(betti_curve) + ggplot2::ggtitle('Betti curves for data')
```

The tibble have two values in dimentions corresponding to the first and second homology dimention. The radius of the filtration of the complex varies from 0 to approximatelly 0.3. The value of the curve is reflected on the y-axis. As expected, the dimension zero (connected components) vanish to zero as the radius grows. The 1-dimension loops grows and then decay.

# Contact me

Please use the issues of the repository or send me a private message with GitHub or LinkedIn <https://www.linkedin.com/in/gonzalezgouveia/>
